A Bayesian approach to burial networks to infer social changes in northeastern Taiwan during the European colonization period

### 2021-01-07 analysis/paper/paper.Rmd
source 006 file in rmd and cross-reference figures

<!-- burial data and the definition of tie formation  --> 
We used burial goods in common to construct ties for representing a flow of good between burials. Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials. Gold-foil beads are in levels of high (> 3), upper-middle (= 3), lower-middle (= 2), and low (= 1); carnelian beads are in levels of high (> 6), upper-middle (= 4-6), lower-middle (= 3), and low (= 1-2); glass beads are in levels of high (> 6), upper-middle (= 3-6), lower-middle (= 2), and low (= 1). That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common (i.e. presence or absence). Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and burial value index. Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records. The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>= 29, account for 10%), above average (= 12-28), and below average (< 12), as an index of wealth. The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.

@@

<!--posterior stats for pre-E-->
The networks constructed from observed burial data show that the network after the European presence has higher node degrees in general (Figure \@ref(fig:network-diagrams)). We compared estimated statistics from the posterior distributions to examine their differences in structure of the simulated networks (Table XX). For the pre-European network simulations, two nodal covariates, age-homophily and ritual-homophily, have significant effects on the formation of relations between burials. The age homophily shows a negative effect that indicates burials of the same age tend to not share similar prestige goods, while the ritual-homophily demonstrates the positive effect that means burials with ritual pottery tend to form relations. Despite positive values for some covariates, such as sex-homophily and burial value-homophily, they do not show a significant tendency to form relations duo to value of zero in the confidence intervals. Similarly, the dyadic covariate, physical distance, shows positive effects but not significantly, indicating that physical proximity does not affect relations. For the endogenous network effects, transitivity (gwesp) present significant positive effect, while popularity (gwdegree) demonstrates negative effect. The positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, indicative of the presence of multiple corporate groups sharing burial goods in common. The negative popularity shows there is a tendency toward decentralization that reflects most burials have similar number of ties without any prominent ones. This might imply that individuals have equal access to trade goods in terms of the flow of goods.

```{r figure-network-diagrams, fig.cap = "Map illustrating the location of burials by periods at the central excavation area of Kiwulan (each square is 4 X 4 meters). The gray dots are post holes."}
knitr::include_graphics(here::here("analysis", "figures", "005-network-diagrams.png"))
```

<!--posterior stats for for post-E and their comparison-->
For the network after the arrival of Europeans the nodal covariates of age-homophily, sex-homophily, ritual-homophily, and the dyadic covariate of physical distance all show positive effects, while the burial value-homophily has a negative effect. However, there are no significant effects for these nodal and dyadic covariates. We also found that the burials in the same wealth level tend not to form relations. Similar to the pre-European network, the endogenous network variable, transitivity (gwesp) demonstrates a significant positive effect. In contrast to the pre-European network, the popularity (gwdegree) shows a significant positive effect. This means there is a tendency toward centralization that reflects a limited number of burials have many more ties than others. This implies that a few central individuals may have better access to trade goods and developed personal networks through wealth accumulation. In general, the post-European network has a smaller transitivity effect and a positive popularity effect than the pre-European network. This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence. Both posterior estimates present symmetric distributions, with the posterior means located close to posterior medians (Table XX).

@@

<!--assessment of MCMC analysis and BGOF--> 
Bayesian goodness of fit (GOF) diagnostics plots show that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e. the number of edges between node pairs in a shortest path [@Hunter2008]) and the degree distribution. For edgewise shared partner (gwesp) distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it. This also suggests that our observed network statistics do not follow a normal distribution, especially for the network after the arrival of the Europeans that presents a bimodal pattern. We compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means. Figure \@ref(fig:xxxx) shows that modeled values from the pre-European network are closer to the observed values overall, compared to the values from the post-European network. For the variance of degree, the values from the modeled post-European distribution are closer to the observed distribution. In general, this demonstrates a better fit for the pre-European network. However, it should be noted that the differences between the statistics of two models are relatively small, which still allows direct comparison of two networks.

>>>>

<!-- burial data and the definition of tie formation  --> 
We used burial goods in common to construct ties for representing a flow of good between burials. Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials. Gold-foil beads are in levels of high (> 3), upper-middle (= 3), lower-middle (= 2), and low (= 1); carnelian beads are in levels of high (> 6), upper-middle (= 4-6), lower-middle (= 3), and low (= 1-2); glass beads are in levels of high (> 6), upper-middle (= 3-6), lower-middle (= 2), and low (= 1). That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common (i.e. presence or absence). Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and burial value index. Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records. The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>= 29, account for 10%), above average (= 12-28), and below average (< 12), as an index of wealth. The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship. The networks constructed based on this principles show that the network after the European presence has higher node degrees in general (Figure \@ref(fig:figure-network-diagrams)).

```{r figure-network-diagrams, fig.cap = "A: Burial network before the European arrival, B: Burial network after the European arrival. The size of the nodes is proportional to the node degrees that means the number of connections to a node. The thickness of tie represents the number of goods in common between nodes."}
knitr::include_graphics(here::here("analysis", "figures", "005-network-diagrams.png"))
```

@@

```{r posterior-distribution, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("006-bergm-distribution-stats.R"))}
suppressWarnings(source(here::here("analysis", "paper", "006-bergm-distribution-stats.R")))
```

<!--posterior stats for pre-E-->
We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table XX; Figure \@ref(fig:figure-posterior-distribution)). For the pre-European network simulations, two nodal covariates, age-homophily and ritual-homophily, have significant effects on the formation of relations between burials. The age homophily shows a negative effect that indicates burials of the same age tend to not share similar prestige goods, while the ritual-homophily demonstrates the positive effect that means burials with ritual pottery tend to form relations. Despite positive values for some covariates, such as sex-homophily and burial value-homophily, they do not show a significant tendency to form relations duo to value of zero in the confidence intervals. Similarly, the dyadic covariate, physical distance, shows positive effects but not significantly, indicating that physical proximity does not affect relations. For the endogenous network effects, transitivity (gwesp) present significant positive effect, while popularity (gwdegree) demonstrates negative effect. The positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, indicative of the presence of multiple corporate groups sharing burial goods in common. The negative popularity shows there is a tendency toward decentralization that reflects most burials have similar number of ties without any prominent ones. This might imply that individuals have equal access to trade goods in terms of the flow of goods.

```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, GWESP, GWDeg, and distance-dyadcov by phases. The pre-European group presents remarkedly larger values for the GWESP, but lower vlaues for the edges and GWDeg parameters comapred to the post-Eruoepan group. The distance parameters overlap between two groups."}
knitr::include_graphics(here::here("analysis", "figures", "006-posterior-distribution.png"))
```

<!--posterior stats for for post-E and their comparison-->
For the network after the arrival of Europeans the nodal covariates of age-homophily, sex-homophily, ritual-homophily, and the dyadic covariate of physical distance all show positive effects, while the burial value-homophily has a negative effect. However, there are no significant effects for these nodal and dyadic covariates. We also found that the burials in the same wealth level tend not to form relations. Similar to the pre-European network, the endogenous network variable, transitivity (gwesp) demonstrates a significant positive effect. In contrast to the pre-European network, the popularity (gwdegree) shows a significant positive effect. This means there is a tendency toward centralization that reflects a limited number of burials have many more ties than others. This implies that a few central individuals may have better access to trade goods and developed personal networks through wealth accumulation. In general, the post-European network has a smaller transitivity effect and a positive popularity effect than the pre-European network . This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence. Both posterior estimates present symmetric distributions, with the posterior means located close to posterior medians (Table XX).

@@

<!--assessment of MCMC analysis and BGOF--> 
Bayesian goodness of fit (GOF) diagnostics plots show that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e. the number of edges between node pairs in a shortest path [@Hunter2008]) and the degree distribution. For edgewise shared partner (gwesp) distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it. This also suggests that our observed network statistics do not follow a normal distribution, especially for the network after the arrival of the Europeans that presents a bimodal pattern. We compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means. Figure \@ref(fig:figure-distribution-moments) shows that modeled values from the pre-European network are closer to the observed values overall, compared to the values from the post-European network. For the variance of degree, the values from the modeled post-European distribution are closer to the observed distribution. In general, this demonstrates a better fit for the pre-European network. However, it should be noted that the differences between the statistics of two models are relatively small, which still allows direct comparison of two networks.

```{r figure-distribution-moments, fig.cap = "Comparison of statistics calculated on the observed and simulated GOF distributions for models between phase groups "}
knitr::include_graphics(here::here("analysis", "figures", "006-distribution-moments.png"))
```
